---
title: "Blog Post 1 - Wine Blog"
author: "Derek Shambo"
date: "November 15, 2018"
output: rmarkdown::github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
#this code will check to see if the package pacman needs to be installed, and if it does, will install it and load it. pacman comes with a function (p_load) that does something similar, so all other packages will be loaded with p_load
if (!require("pacman")) install.packages("pacman") 
p_load(tidyverse,dplyr,factoextra)
```

The analysis conducted will utilize clustering. The goal here will be to see if there is some natural segmentation in the wines. This could be utilized to help someone understand find other wines that they enjoy, or wines that they should avoid. It may also be useful in helping wine makers understand what chemical properties create certain flavor profiles.

Data input and exploration:
The data being analyzed is different properties of wine. It is the results of chemical analysis conducted on the wines. All of the wines come from the same region in Italy, and from 3 different cultivars in that region.

```{r}
#data load in. Col names taken from uci website
wine_data <- read_csv("https://archive.ics.uci.edu/ml/machine-learning-databases/wine/wine.data",
                      col_names = c("Col 1",
                                    "Alcohol",
                                    "Malic acid",
                                    "Ash",
                                    "Alcalinity of ash",
                                    "Magnesium",
                                    "Total phenols",
                                    "Flavanoids",
                                    "Nonflavanoid phenols",
                                    "Proanthocyanins",
                                    "Color intensity",
                                    "Hue",
                                    "OD280/OD315 of diluted wines",
                                    "Proline"))

#col 1 is a class item, but there isn't documentation on what those classes actually represent, so this column will be removed
wine_data$`Col 1` <- NULL

#check for NAs
sum(is.na(wine_data)) #There are no NAs present in the data!

#lets take a look at the data
summary(wine_data)
```

Luckily our data didn't require a lot of preparation, and we are ready to dive into the analysis. As discussed before, we will be using cluster analysis to try to get a better understanding of what is in our wine data. 

Cluster Creation:
The first thing we need to do is create our clusters. For this we will scale our data. This insures that all variables are utilized equally in the cluster analysis. We will then create our clusters using hierarchical clustering. Using a dendogram, we can visualize our clusters and see how our data seperates.

```{r}
#The first thing we want to do is scale the data
wine_scale <- as.matrix(scale(wine_data))

d <- dist(wine_scale, method = "euclidean") # distance matrix
fit <- hclust(d, method="ward.D2")
plot(fit) # display dendogram
#as we can see, there are 3 clusters that are evident
abline(h=15)

rect.hclust(fit, k=3, border="red")
#the boxes displayed show our three clusters
```

Analysis of clusters:
We will now "cut" our dendogram at the specified point to get our 3 clusters. These clusters will be bound with our data, allowing us to analysis what those clusters actually say. We will then look at the mean value for each of our variables for each cluster. This will tell us what is unique about each cluster.
```{r}
#first, we will cut our dendogram at 3 clusters
cut <- cutree(fit, k = 3)

#next, we will bind our data with that cut to know which wines belong wo which clusters
wine_clust <- cbind(wine_data, cut)

#this will allow us to visualize the clusters
fviz_cluster(list(data = wine_clust[-14], cluster = cut))

(sum_clusts <- wine_clust %>%
     group_by(cut) %>%
     summarise_all(funs(mean)))
#we can now view sum_clusts to see what each cluster looks like, and how they compare to each other

```

Wines in Cluster 1:
High alcohol content
Average amount of Malic Acid
High Ash
Low Alcalinity of ash*
High Magnesium
High total phenols
High Flavanoids*
Low Nonflavanoi phenols
High Proanthocyanins
Average color intensity
High hue
High OD280/OD315 of diluted wines
High Proline*

Wines in Cluster 2:
Low alcohol content
Average Malic acid
Low ash*
Average alcalinity of ash
Low Magnesium
Average Flavanoids
Average Nonflavanoid phenols
Average Proanthocyanins
Low color intensity*
Average hue
Average OD280/OD315 of diluted wines
Low Proline

Wines in Cluster 3:
Average alcohol content
High Malic acid*
High ash
High alcalinity of ash
Medium Magnesium
Low Flavanoids*
High Nonflavanoid phenols
Low Proanthocyanins*
High color intensity
Low hue
Low OD280/OD315 of diluted wines*
medium-low Proline

*significantly lower or higher than the other two clusters

Looking at our dendogram and cluster visualization, there was a pretty clear split for these three clusters. 
